---
title: "Oil palm SOC and leaf litter decomposition"
author: "Marcus Schmidt"
date: "8 November 2019"
output: word_document
---

### Background
In the following course, we are looking at two processes in oil palm plantations:
1) Soil organic carbon (SOC) remaining after deforestation and 2) Decomposition of oil palm leafs

### Part 1 - SOC remaining
This data is part of a study on deforestation and land luse conversion in Indonesia, Cameroon and Peru. You can find the whole paper here: https://www.pnas.org/content/112/32/9956

Let's fist load our library. It is the tidyverse library, which includes useful tools for data handling and visualization, i.e. the ggplot 2 package.
```{r libraries, echo = T, results='hide', message=F, warning=F}
library("tidyverse")
```

Let's now read in our data and look at the upper part:

```{r data-read-in, echo = T, results = 'hide'}
DATA=read.csv("data/van-straaten-soc.csv",h=T)
head(DATA)
```

I like to shorten our variables a little bit, it will be easier to type then:

```{r shortening, echo = T}
DATA$t<-DATA$Time.since.Deforestation
DATA$soc_pc<-DATA$SOC_prop_top10_C2
```

Next, I apply a filter to exclude observations that don't state the country. We also want only oil palm observations and lastly, clay content shouldn't change more than 20%.

```{r filtering, echo = T}
DATA<-DATA%>%filter(Country!="")%>%filter(Landuse == "Oil palm")
DATA<-DATA%>%filter(abs(d_Clay_50to100)<20)
```

Now, for some of the following coding, it's easier to give a shorter name to some variables. We choose x and y because this we will have on our axes:

```{r shortening-variables, echo = T}
y <- DATA$soc_pc #
x <- DATA$t
```

Let's plot this in a very basic form, do the fitting and then add the line:
```{r basic-plot, echo = T}
xx <- seq(0,605, length=1500)
plot(y~x)

fit2=nls(y~a+(100-a)*exp(-k*x),start=list(a=60, k=0.15), algorithm="port")

lines(xx, predict(fit2, data.frame(x=xx)), col="red", lwd=2, lty=1)
summary(fit2)
```

The plot will look much nicer in ggplot2, however:
```{r nicer-plotting, echo = T}
g2 <- ggplot(DATA,aes(t,soc_pc, shape = Country, colour = Country)) +
    geom_hline(yintercept=100) +
  xlim(0,50) +
  ylim(0,150) +
  geom_point(show.legend = T, size = 3) +
  ggtitle("Change in organic carbon") +
  xlab("Years since deforestation") +
  ylab("Percent carbon remaining") +
  theme_bw()
g2
```


Finally, we will add the curve to our ggplot:
```{r final plotting, echo = T}
func3<- function(x) {60.4+(100-60.4)*exp(-0.1296*x)}

g2 +
  stat_function(fun = func3, xlim=c(0,400), col = "blue", size = 1) 
```




### Part 2 - Decomposition
This data was collected in Indonesia where oil palm leafs are put in between oil palm rows and left there to decompose.

I am first going to load my data from the .txt file I created and look at the first 5 observations too see what's going on:
```{r data-loading, echo = T, results = 'hide'}
DATA=read.table("data/leaf-decomp-data.txt",h=T)
head(DATA)
```

You may also want to try
```{r data-viewing, echo = T, results = 'hide', eval = F}
View(DATA)
```
to see the whole data set.

We now want to look at the first soil type (Clay Acrisol) at the first site, so we create a subset like this, using the pipe (%>%) from the tidyverse package. Then we check if this really worked.

```{r filtering-2, echo = T}
DATA<-DATA%>%filter(Plot == "C1")
unique(DATA$Plot)
```

The next part is key! We are plotting our Data - days and mass, fit an exponential cuve and retrieve k, our decomposition rate:


```{r fitting-and-plotting, echo = T}
xx <- seq(0,605, length=1500)
plot(DATA$Mass_g 
     ~ DATA$d)
y <- DATA$Mass_g
x <- DATA$d
fit <- nls(y ~ (50.81 * 2.718^(-k * x)), start=list(k=0.00001))
lines(xx, predict(fit, data.frame(x=xx)), col="red", lwd=2, lty=1)
summary(fit)
```

In the output above, we are given the k and some information on how well it fits the observed data.

We made the plot using the basic plot package in R. However, in order to get a much nicer plot, I recommend using the ggplot 2 package. It's already included in the tidyverse library we installed earlier. 

We first define the function including k, which we get from above and define what data we want to use:

```{r function-defining, echo = T, }
func1<- function(x) {50.81 * 2.718^(-0.0052707 * x)}
dat1 = data.frame(x = DATA$d, y = DATA$Mass_g)
```

Now we define the plot:
```{r gg-plot-code, echo = T, results = 'hide', eval = F}
g1 <- ggplot(dat1,aes(x,y)) +
  xlim(0,400) +
  ylim(0,60) +
  geom_point(show.legend = FALSE) +
  ggtitle("Leaf mass decline on Clay Aricsol 1") +
  xlab("days") +
  ylab("remaining mass in g") +
  stat_function(fun = func1, xlim=c(0,400), col = "red")
g1
```

And this is what we get:
```{r gg-plot-output, echo = F}
g1 <- ggplot(dat1,aes(x,y)) +
  xlim(0,400) +
  ylim(0,60) +
  geom_point(show.legend = FALSE) +
  ggtitle("Leaf mass decline on Clay Aricsol 1") +
  xlab("days") +
  ylab("remaining mass in g") +
  stat_function(fun = func1, xlim=c(0,400), col = "red") +
  theme_bw()
g1
```

To save your plot in a folder called figs within your project, we can use the cowplot package.
```{r saving, echo = T, results = 'hide', message=F, warning=F}
library(cowplot)
ggsave("figs/oil-palm-decomp.png", plot = g1, width = 10, height = 10)
```

