---
title: "Oil palm SOC and leaf litter decomposition"
author: "Marcus Schmidt"
date: "8 November 2019"
output: github_document
---

```{r include = F}
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)
```


### Background
In the following course, we are looking at two processes in oil palm plantations:
1) Soil organic carbon (SOC) remaining after deforestation and 2) Decomposition of oil palm leafs

### Part 1 - SOC remaining
This data is part of a study on deforestation and land luse conversion in Indonesia, Cameroon and Peru. You can find the whole paper here: https://www.pnas.org/content/112/32/9956

Let's fist load our library. It is the tidyverse library, which includes useful tools for data handling and visualization, i.e. the ggplot 2 package.
```{r libraries, echo = T, results='hide', message=F, warning=F}
library("tidyverse")
```

Let's now read in our data and look at the upper part:

```{r data-read-in, echo = T, results = 'hide'}
DATA=read.csv("data/van-straaten-soc.csv",h=T)
head(DATA)
```

I like to shorten our variables a little bit, it will be easier to type then:

```{r shortening, echo = T}
DATA$t<-DATA$Time.since.Deforestation
DATA$soc_pc<-DATA$SOC_prop_top10_C2
```

Next, I apply a filter to exclude observations that don't state the country. We also want only oil palm observations and lastly, clay content shouldn't change more than 20%.

```{r filtering, echo = T}
DATA<-DATA%>%filter(Country!="")%>%filter(Landuse == "Oil palm")
DATA<-DATA%>%filter(abs(d_Clay_50to100)<20)
```

Now, for some of the following coding, it's easier to give a shorter name to some variables. We choose x and y because this we will have on our axes:

```{r shortening-variables, echo = T}
y <- DATA$soc_pc #
x <- DATA$t
```

Let's plot this in a very basic form, do the fitting and then add the line:
```{r basic-plot, echo = T}
xx <- seq(0,50, length=500)
plot(y~x)

fit2=nls(y~a+(100-a)*exp(-k*x),start=list(a=60, k=0.15), algorithm="port")

lines(xx, predict(fit2, data.frame(x=xx)), col="red", lwd=2, lty=1)
summary(fit2)
```

This is a nice summary, but how do we extract k, for example?
```{r extracting K, echo = T}
summary(fit2)$coefficients
k <- summary(fit2)$coefficients[2,1]
k
```

#### Calculating turnover time

k can be though of as the fraction of organic carbon that is lost every year. 
We can therefore calculate turnover time by 1/k. 1 is the whole of the carbon at
the beginning so how many k's (fractions per year) do we need to lose until everything
is gone/turned over?

```{r calculating turnover time}
turnover_time <-  1/k
turnover_time
print(paste0("Our turnover time is: ",round(turnover_time,2)," years"))

```



The plot will look much nicer in ggplot2, however:
```{r nicer-plotting, echo = T}
g2 <- ggplot(DATA,aes(t,soc_pc, shape = Country, colour = Country)) +
    geom_hline(yintercept=100) +
  xlim(0,50) +
  ylim(0,150) +
  geom_point(show.legend = T, size = 3) +
  ggtitle("Change in organic carbon") +
  xlab("Years since deforestation") +
  ylab("Percent carbon remaining") +
  theme_bw()
g2
```


Finally, we will add the curve to our ggplot:
```{r final plotting, echo = T}
func3<- function(x) {60.4+(100-60.4)*exp(-0.1296*x)}

g2 +
  stat_function(fun = func3, xlim=c(0,400), col = "blue", size = 1) 
```


#### Exercise (to be added to course folder):

Add your name and the turnover time into the plot using annotate(). Create a PDF file including your plot and the code you used. Add this to the course folder.



### Part 2a - Decomposition of Clay Acrisol 1
This data was collected in Indonesia where oil palm leafs are put in between oil palm rows and left there to decompose.

I am first going to load my data from the .txt file I created and look at the first 5 observations too see what's going on:
```{r data-loading, echo = T, results = 'hide'}
DATA=read.table("data/leaf-decomp-data.txt",h=T)
head(DATA)
```

You may also want to try
```{r data-viewing, echo = T, results = 'hide', eval = F}
View(DATA)
```
to see the whole data set.

We now want to look at the first soil type (Clay Acrisol) at the first site, so we create a subset like this, using the pipe (%>%) from the tidyverse package. Then we check if this really worked.

```{r filtering-2, echo = T}
DATA<-DATA%>%filter(Plot == "C1")
unique(DATA$Plot)
```

The next part is key! We are plotting our Data - days and mass, fit an exponential cuve and retrieve k, our decomposition rate:


```{r fitting-and-plotting, echo = T}
xx <- seq(0,605, length=1500)
plot(DATA$Mass_g 
     ~ DATA$d)
y <- DATA$Mass_g
x <- DATA$d
fit <- nls(y ~ (50.81 * 2.718^(-k * x)), start=list(k=0.00001))
lines(xx, predict(fit, data.frame(x=xx)), col="red", lwd=2, lty=1)
summary(fit)
```

In the output above, we are given the k and some information on how well it fits the observed data.

We made the plot using the basic plot package in R. However, in order to get a much nicer plot, I recommend using the ggplot 2 package. It's already included in the tidyverse library we installed earlier. 

We first define the function including k, which we get from above and define what data we want to use:

```{r function-defining, echo = T, }
func1<- function(x) {50.81 * 2.718^(-0.0052707 * x)}
dat1 = data.frame(x = DATA$d, y = DATA$Mass_g)
```

Now we define the plot:
```{r gg-plot-code, echo = T, results = 'hide', eval = F}
g1 <- ggplot(dat1,aes(x,y)) +
  xlim(0,400) +
  ylim(0,60) +
  geom_point(show.legend = FALSE) +
  ggtitle("Leaf mass decline on Clay Aricsol 1") +
  xlab("days") +
  ylab("remaining mass in g") +
  stat_function(fun = func1, xlim=c(0,400), col = "red")
g1
```

And this is what we get:
```{r gg-plot-output, echo = F}
g1 <- ggplot(dat1,aes(x,y)) +
  xlim(0,400) +
  ylim(0,60) +
  geom_point(show.legend = FALSE) +
  ggtitle("Leaf mass decline on Clay Aricsol 1") +
  xlab("days") +
  ylab("remaining mass in g") +
  stat_function(fun = func1, xlim=c(0,400), col = "red") +
  theme_bw()
g1
```

To save your plot in a folder called figs within your project, we can use the cowplot package.
```{r saving, echo = T, results = 'hide', message=F, warning=F}
library(cowplot)
ggsave("figs/oil-palm-decomp.png", plot = g1, width = 10, height = 10)
```

### Part 2b - Decomposition representing the whole study area
We have previously looked at one of our plots but that does not represent our whole study area. Also, we want to get rid of our subplots because only our plots are true replicates.

So first we read in our data and only keep the first 5 columns:

```{r ch1, echo = T}
DATA=read.table("data/leaf-decomp-data.txt",h=T)
names(DATA)
selection<-names(DATA)[1:5]
DATA<-DATA%>%select(selection)
head(DATA)
```


We then group and summarize our data - assigning a mean for each plot and day. For our example this means that we do not anymore distinguish by subplot.

```{r ch2, echo = T, , warning=F}
DATA<-DATA%>%group_by(day, Time, Plot)%>%summarize(plot_mass_g = mean(Mass_g))
DATA<-data.frame(DATA)
head(DATA)
```

Now we define our x, our y and the sequence for our fitting

```{r ch3, echo =T}
x<-DATA$day
y<-DATA$plot_mass_g
xx <- seq(0,400, length=800)
```

Next is the fitting itself. Again, we take 50g as the basis from which decomposition starts.

```{r ch3b, echo=T, results = "hide"}
fit <- nls(y ~ (50 * exp(-k * x)), start=list(k=0.00001))
summary(fit)

```


To have it handy later, we are saving our k of the fit

```{r ch4, echo = T}
k<-0.0048034
```

You can now do the simple plot like this, but I will not show it here:
```{r ch5, echo = T}
# plot(x,y)
# lines(xx, predict(fit, data.frame(x=xx)), col="red", lwd=2, lty=1)
```

Instead, let's do it with ggplot:

```{r ch6, echo = T}
func1<-func1<- function(x) {50 * exp(-k * x)}

g1 <- ggplot(,aes(x,y)) +
  xlim(0,400) +
  ylim(0,60) +
  geom_point(show.legend = FALSE) +
  ggtitle("Leaf mass decline on all plots") +
  xlab("days") +
  ylab("remaining mass in g") +
  stat_function(fun = func1, xlim=c(0,400), col = "red") 
g1
```


This looks nice alreay, but to give a more common representation, let's make it a line that includes the plot means and standard errors for each month:

```{r ch7, echo = T}
DATA<-DATA%>%group_by(day,Time)%>%summarize(mean_mass_g = mean(plot_mass_g), se_g = sd(plot_mass_g/sqrt(4)))
head(DATA)
x<-DATA$day
y<-DATA$mean_mass_g
z<-DATA$se_g # for error visability

library(sciplot)

g2 <- ggplot(,aes(x,y)) +
  xlim(0,400) +
  ylim(0,60) +
  geom_line(show.legend = FALSE) +
  ggtitle("Leaf mass decline for area") +
  xlab("days") +
  ylab("remaining mass in g") +
  stat_function(fun = func1, xlim=c(0,400), col = "red")  +
  geom_errorbar(aes(ymin=y-z, ymax=y+z), colour="black", width=.1) 
g2


```





